FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-08-30 2018EA000388R_Answers_Reviewers.txt

Reviewer 1

> Page 4, the footnote 8 should rather be psa.esa.int (which is indeed
> redirected to the addressed mentioned). In the future, the psa.esa.int
> will certainly remain.

Fixed

> L278 -> headers

Fixed

-------------------------------------------------------------------------------

Reviewer 2

We want to thank the anonymous reviewer for his thoughtful comments.
Sections 2.1 and 3 have been enriched following his suggestions.

> 1. The FITS tool ecosystem - FITS has been around for a long time, and has a
> robust user base. Accordingly, the FITS tool ecosystem is very large and well
> established. This seems like a prime motivator for moving toward FITS.
> This paper notes on lines 72-73 that there is a "large number of open libraries
> and software tools," but does not say much more. It would be useful to note
> that FITS tools exist in numerous languages, including, but not limited to, C,
> C++, IDL, Java, MATLAB, Fortran, R, & Python.

Section 2.1 has been expanded (see paragraph starting at L68 in trackchanges.pdf),
references to available tools have been added in footnotes.

> 2. This leads to my second point, which is that section 3 needs more explanation.
> At minimum, there should be an introduction to the section that describes why
> these specific tools are featured. If the paper provides more description up front
> of the current FITS tool ecosystem, this section 3 could then be contextualized as
> illustrating how the FITS tool ecosystem is still expanding, and/or that for the
> planetary data community to fully take advantage of FITS, more expansion is possible,
> and already in progress.

A small introduction to Section 3 has been added, in particular clarifying, we hope,
the choice of the described tools.

> 3. A bigger picture question that this paper should address is, "why hasn't this
> already happened?" Again, FITS is over 30 years old. What barriers or disincentives
> have been or currently are in place for the planetary data community preventing or
> slowing a move toward FITS? The paper should then describe why this proposed scheme
> reduces or bypasses the relevant barriers, or provides incentives to make the move.
> Is it just tools, as I note above? Or are there community reasons, e.g. that
> astronomical and planetary communities have been disjoint to such an extent that
> languages, standards, etc have never dovetailed? What are the authors' perspectives
> on such barriers/disincentives, and how the present paper smooths over or bypasses them?

Section 2.1 has been expanded (see paragraph starting at L86 in trackchanges.pdf).
We hope this syntetic answer will be enough to clarify our point of view.
A more detailed discussion have been omitted due to overall length.

> 4. Editorial comment - the paper is heavy on acronyms, some of which are not spelled
> out at all, e.g. VRT, some of which are only spelled out in the titles of papers in the
> bibliography but not in the text (VESPA, CRISM, HiRISE).
> All acronyms should be spelled out, and/or a table of acronyms should be added.

Fixed.

### 2018-08-30 ess_marmo.tex
Last edits and answers to the reviewers.

\subsubsection{Visualization of Global and Projected Maps}
\label{sssec:maps}
High level products are meant to be used for visual inspection or GIS analysis.
They are distributed as projected rasters, often in
GeoTIFF\footnote{http://geotiff.osgeo.org/}
format, as GeoTIFF make them GIS ready.
Given some new keywords introduced in FITS and WCS definitions adapted to planetary
surface, FITS and GeoTIFF are completely interchangeable.
As an example we have downloaded a Mercury polar map from the United States Geological Survey (USGS)
PDS Annex\footnote{https://astrogeology.usgs.gov/search} and converted it in FITS.
The FITS image has been cropped in two smaller versions using the open source software
SWarp\footnote{https://www.astromatic.net/software/swarp} \citep{swarp}: SWarp was able to
automatically manage WCS planetary information.
We used ds9 to display the three FITS images and were able to correctly align them.
Linear coordinates in meters are visualized (see figure \ref{fig:ds9qgis}, left side)
as in QGIS, a popular open source GIS software, thanks to the alternate coordinate
description keywords in the header.
\begin{figure}[ht!]
\centerline{\includegraphics[height=17pc]{ess_marmo2} \includegraphics[height=17pc]{ess_marmo1}}
\caption{left: Mercury South Pole FITS images displayed in ds9, alternate coordinates in meters
are displayed;
right: GeoTIFF raster (displayed in red) and the two virtual haeders
linked to FITS rasters (displayed in green and blue); the three layers
are perfectly superposed.
Mercury image credits: MESSENGER Team, Arizona State University, Johns Hopkins Applied
Physics Laboratory, Carnegie Science}
\label{fig:ds9qgis}
\end{figure}

>>>>

\subsubsection{Visualization of Global and Projected Maps}
\label{sssec:maps}
High level products are meant to be used for visual inspection or GIS analysis.
They are distributed as projected rasters, often in
GeoTIFF\footnote{http://geotiff.osgeo.org/}
format, as GeoTIFF make them GIS ready.
Given some new keywords introduced in FITS and WCS definitions adapted to planetary
surface, FITS and GeoTIFF are completely interchangeable.
As an example we have downloaded a Mercury polar map from the United States Geological Survey (USGS)
PDS Annex\footnote{https://astrogeology.usgs.gov/search} and converted it in FITS.
The FITS image has been cropped in two smaller versions using the open source software
SWarp\footnote{https://www.astromatic.net/software/swarp} \citep{swarp}: SWarp was able to
automatically manage WCS planetary information.
We used ds9 to display the three FITS images and were able to correctly align them.
Linear coordinates in meters are visualized (see figure \ref{fig:ds9qgis}, left side)
as in QGIS, a popular open source GIS software, thanks to the alternate coordinate
description.
\begin{figure}[ht!]
\centerline{\includegraphics[height=17pc]{ess_marmo2} \includegraphics[height=17pc]{ess_marmo1}}
\caption{left: Mercury South Pole FITS images displayed in ds9, alternate coordinates in meters
are displayed;
right: GeoTIFF raster (displayed in red) and the two virtual headers
linked to FITS rasters (displayed in green and blue); the three layers
are perfectly superposed.
Mercury image credits: MESSENGER Team, Arizona State University, Johns Hopkins Applied
Physics Laboratory, Carnegie Science}
\label{fig:ds9qgis}
\end{figure}
